From khudian@manchester.ac.uk Mon Oct  9 13:37:51 2017
Date: Mon, 9 Oct 2017 13:37:51 +0100
From: Hovhannes Khudaverdian <khudian@manchester.ac.uk>
To: veronica.kelsey@student.manchester.ac.uk
Cc: mark.coleman@manchester.ac.uk
Subject: Galois Group lectures 22 November?


            Dear Veronica
   I am terribly sorry, but
   you did not answer my previous letter.

   Do I have to understand that the day 22 November Wednesday
   is fixed, or no?

    Thank you

                     HMK


                             Dr Hovhannes Khudaverdian
                         Senior Lecturer in Pure Mathematics
                              School of Mathematics
                           The University of Manchester
                       Oxford Road, Manchester  M13 9PL, UK
                         e-mail:khudian@manchester.ac.uk
                          tel 00(44)-(0)-161-200-36-82

---------- Forwarded message ----------
Date: Wed, 4 Oct 2017 08:09:28 +0100
From: khudian@manchester.ac.uk
To: veronica.kelsey@student.manchester.ac.uk
Cc: mark.coleman@manchester.ac.uk
Subject: Re: Galois Group lectures


         Yes, I am able.
    I will send you the abstract a week before.

                   HMK

                              Dr Hovhannes Khudaverdian
                          Senior Lecturer in Pure Mathematics
                               School of Mathematics
                            The University of Manchester
                        Oxford Road, Manchester  M13 9PL, UK
                          e-mail:khudian@manchester.ac.uk
                           tel 00(44)-(0)-161-200-36-82

On Tue, 3 Oct 2017, Veronica Kelsey wrote:

>  Hi Dr Khudaverdian,
>  Sorry for the delay in replying! It would be great
> if you could do a Galois Group talk, and your sketch of an
> abstract sounds really interesting.
> I’ve currently got one organised with Dr Tom Kempton for November 8th,
> and I wasn’t planning on having them too close together.
> Would you be able to do the 22nd of November instead?
>  Thanks,
>  Veronica
> 
> >  On 26 Sep 2017, at 5:27 pm, Hovhannes Khudaverdian 
> >  <khudian@manchester.ac.uk> wrote:
> > 
> > 
> >                Dear Veronica,
> >     I am Hovhannes Khudaverdian. Dr Coleman told me that you are organiser 
> >     of Galois group  lectures.  This is great work!  Thank you.
> >     I would like to have the Galois lecture this semester
> >     The title of the lecture:
> >     "Locally Euclidean geometries and hyperbolic geometry"
> >     Here is the plan of the lecture.
> >  (It is not the abstract, this is just the sketch of the abstract)
> >  ********************************************************************
> >    In the first part of talk we will discuss so called locally Euclidean
> >  two dimensional geometries:
> >  Ant-mathematician  living on cylindre, or on the torus will not 
> >  distinguish the geometry of the surface from the Euclidean Geometry on 
> >  small distances. For small triangles, the sum of the angles will be
> >  $\pi$, the Pythagorean Theorem will be the same. One can classify all the 
> >  surfaces  which have locally Euclidean geometry. They are:
> >  surface of cylinder, surface of torus, and also surface of twisted 
> >  cylinder (Mobius band) and so called Klein bottle.
> >    In the second part of the talk we will consider the set of all these
> >  geometries, and will show that this set can be described by the points
> >  of Lobachevsky (hyperbolic) plane . Thus we will come to
> >  Lobachevsky (hyperbolic) geometry.
> >  ***********************************************************************
> >    I repeat again: this is not the abstract, this is just the sketch of 
> >  abstract. In particular it has to be written better (and
> >  English has to be corrected also!).
> >    Now  about timing of the lecture: I will prefer to do
> >  it in the beginning of Novemeber, but this question may be has to be 
> >  discussed.  What are your intentions? May be we have to  meet to discuss 
> >  this question.
> >     Tomorrow I will be all the day: I have classes from 9-10,
> >  then  I will wait meetings with tutees from 10.30--11.30
> >  (we can speak this time but we have to interrupt if any of tutees will 
> >  come),
> >  then I will be available after 3pm  till late,
> >     On Thursday I will not be available.
> >     On Friday we may meet around 1pm (I will be busy after 2pm)
> >     Sure we may meet next week on Monday (after 1pm)
> >  or on Tuesday (after 10am)....
> > 
> >                    Best regars
> > 
> >                                   HMK
> > 
> > 
> >                             Dr Hovhannes Khudaverdian
> >                         Senior Lecturer in Pure Mathematics
> >                              School of Mathematics
> >                           The University of Manchester
> >                       Oxford Road, Manchester  M13 9PL, UK
> >                         e-mail:khudian@manchester.ac.uk
> >                          tel 00(44)-(0)-161-200-36-82
> 
>