\magnification=1200

   \centerline {One approach to higher order invariants}

 We consider curves---  $x^i=x^i(t)$ in $d$-dimensional
eucledian space. Our goal is to construct local integrands
 over these curves which are

1) reparametrization invariant

2) invariant under global translation  and rotation of curves

3) are full divergence!---i.e. their Euler Lagrange equations are trivial

 If we omit condition (3) the answer is invariant densities--
length: $\sqrt {{dx^i\over dt}{dx^i\over dt}}$,...

But how to construct at least one invariant closed density?
In fact in $d$-dimensional case we restrict ourself by densities
whose rank (the order of derivatives) is not higher than $d$


1) $d=1$ . Answer is trivial:

               $$
        A=  x_t
                         \eqno (1)
                $$

2) $d=2$ Less trivial but I think you know it:
                 $$
 A= {x_ty_{tt}-y_tx_{tt}\over x_t^2+y_t^2}
                               \eqno (2)
                   $$
  One can see easily that (2) is total divergence
 (it is evident in polar coordinates).


3) Now I present you the closed density in $3$-dimensional space:
                      $$
                      A=
                      {
                    \det
                   \pmatrix
                      {
           &x_t &y_t & z_t \cr
           &x_{tt} &y_{tt} &z_{tt} \cr
           &x_{ttt} &y_{ttt} &z_{ttt} \cr
                      }
                      \over
            (x_ty_{tt}-y_{t}x_{tt})^2+
            (y_tz_{tt}-z_{t}y_{tt})^2+
            (z_tx_{tt}-x_{t}z_{tt})^2
                      }
                      \cdot
                \sqrt
              {x_t^2+y_t^2+z_t^2}
                                             \eqno (3)
                        $$

  Do you want to check by straightforward calculation
that (3) is total divergence?
Note please that "razmernostj" in (3) and (2) is $[{1\over sec}]$
I came to (3) using (2)...

In fact I think that in $d$-dimensional space all invariant densities
can be expressed in terms of determinant as in (3) and its
first minors.
For example:

  Let we denote by $x^i_{(n)}$ the $n$-th derivative by $t$ of the
 coordinate $x^i$  in $I\!E^d$  $(i=1,...,d)$.
  We consider  the matrix
                        $$
          \pmatrix
                        {
             &x^1_{(1)}, &x^2_{(1)},  &x^3_{(1)}, ...,&x^d_{(1)} \cr
             &x^1_{(2)}, &x^2_{(2)},  &x^3_{(2)}, ...,&x^d_{(2)} \cr
             &x^1_{(3)}, &x^2_{(3)},  &x^3_{(3)}, ...,&x^d_{(3)} \cr
             &..., &...,  &..., ...,&... \cr
             &..., &...,  &..., ...,&... \cr
             &..., &...,  &..., ...,&... \cr
             &..., &...,  &..., ...,&... \cr
           &x^1_{(d)}, &x^2_{(d)},  &x^3_{(d)}, ...,&x^d_{(d)} \cr
                   }
                          \eqno (4)
                      $$
 We denote
                  $$
 \Delta_1= \sqrt {(x^1_{(1)})^2+(x^2_{(1)})^2+(x^3_{(1)})^2+
      (x^d_{(1)})^2}\,,
                    $$
                    $$
    \Delta_2=\sqrt
                        {
              \left(
                 \det
                             \pmatrix
                                    {
                       &x^1_{(1)}, &x^2_{(1)}\cr
                       &x^1_{(2)}, &x^2_{(2)} \cr
                                        }
                                  \right)^2
                               +
                 \dots+
              \left(
                 \det
                             \pmatrix
                                    {
                       &x^{d-1}_{(1)}, &x^d_{(1)}\cr
                       &x^{d-1}_{(2)}, &x^d_{(2)} \cr
                                        }
                                  \right)^2
                     }
                              $$

                    $$
    \Delta_3=\sqrt
                        {
              \left(
                 \det
                             \pmatrix
                                    {
                       &x^1_{(1)}, &x^2_{(1)},&x^3_{(1)}\cr
                       &x^1_{(2)}, &x^2_{(2)},&x^3_{(2)} \cr
                       &x^1_{(3)}, &x^2_{(3)},&x^3_{(3)} \cr
                                         }
                                  \right)^2
                               +
                 \dots+
              \left(
                 \det
                             \pmatrix
                                    {
                    &x^{d-2}_{(1)}, &x^{d-1}_{(1)}, &x^d_{(1)}\cr
                 &x^{d-2}_{(2)}, &x^{d-1}_{(2)}, &x^d_{(2)} \cr
                 &x^{d-2}_{(3)}, &x^{d-1}_{(3)}, &x^d_{(3)} \cr
                                      }
                                  \right)^2
                     }
                              $$
 e.t.c. $\Delta_d$ is nothing that the determinant of the matrix (4).

We consider in $d$-dimensional space the following invariant densities
 the densities
               $$
 A_1=\Delta_1,\quad  A_2={\Delta_2\over\Delta_1^2},  \quad
         A_3={\Delta_3\cdot\Delta_1\over\Delta_2^2},\dots
           \eqno (5)
            $$
It is easy to see that $A_1$ is nothing but the lenghth of the curve,
and it is divergence in the case $d=1$, $A_2$ is the density of
second rank which coincides with (2)
if $d=2$, $A_3$ is the density of third rank which
coincides with (3) if $d=3$.
Sure one can continue this sequence:
In $d$-dimensional case one can consider
ther sequence of the invariant densities
   $\{A_1,...,A_d\}$ such that
the last one is total divergence...

  Do you like it?
Do you know how I come for example to (3)?
                  \bigskip
         \centerline {\bf Section2 continuation}
                     \medskip
  How one can come to the expression (3)


 We recall at first the essence of the formula (2).
If $L$ is the curve in $I\!E^2$ then the tangent unit
 vector to this curve maps the curve $L$ to the curve in $S^1$:
            $$
             \varphi={\rm arctg}{y_t\over x_t}
              \eqno (2.1)
                $$
 and (2) corresponds to the pull back of the volume form $d\varphi$
  on the $S^1$. In the other word the density (2) is constructed via
 the canonical density on the $S^1$ and the density (2.1)
  {\bf which takes values in the unit vectors}---it is
 the composition of the densities.
                  $$
     A_2={d\over dt} {\rm arctg}{y_t\over x_t}
                $$
 Now bearing in mind this interpretation for (2)
we consider for the curve $L$ in the three-dimensional
space the map analogous to (2.1).---Considering
the unit vector to the every point of the curve
 $L$ we transform this curve to the curve
 on the sphere $S^2$:
                         $$
               \theta=
       {\rm arccos}{z_t\over\sqrt{x_t^2+y_t^2+z_t^2}},\quad
                 \varphi={\rm arctg}{y_t\over x_t}
           \eqno (2.2)
                    $$
 We come to the curve on the $2$-dimensional sphere and we apply
 the formula (2) for this curve. Of course the sphere is not
 euclidean but physical intuition say us that in this case
      for the curves on the $S^2$ the analog of (2) will be something like
                $$
      {d\over dt} {\rm arctg}{\sin\theta \cdot\varphi_t\over \theta_t}
                  $$
Combining this density with previous one we come to
 the "answer"
                $$
  A_3 ={d\over dt} {\rm arctg}{\sin\theta \cdot\varphi_t\over \theta_t}=
     {d\over dt}\left (
        {\rm arctg}{\sin
                 \left(
{\rm arccos}{z_t\over\sqrt{x_t^2+y_t^2+z_t^2}}
           \right)
                       \cdot
        \left(
              {\rm arctg}{y_t\over x_t}
                     \right)_t
                 \over
                  \left(
         {\rm arccos}{z_t\over\sqrt{x_t^2+y_t^2+z_t^2}}
             \right)_t}
                 \right)
            \eqno (2.3)
                      $$

 Can you calcule this? I cannot. It is here where I began to think.
 (All previous constructions, you understand,
 do no need paper in spite of
their "artificiel" complicatness...)

 For considering this tremendous expression I consider first the case
 where  in the given point
                 $$
               z_t\big\vert_{t=t_0}=y_t\big\vert_{t=t_0}=0
               $$
 hence in the monstre (2.3)
                    $$
            \theta(t) \big\vert_{t=t_0}={\pi\over 2},\quad
            \varphi\big\vert_{t=t_0}=0
            $$
  Using the multiple time the fact known in school that
               $$
            \left(f(x)x\right)^\prime_{x=0}=f(0)
                    $$
one comes in 15 minutes to the result
                        $$
       A_2= {y_{tt}z_{ttt}-y_{ttt}z_{tt}\over y_{tt}^2+z_{tt}^2}\quad
                {\rm if}\quad
               z_t\big\vert_{t=t_0}=y_t\big\vert_{t=t_0}=0
                     $$
 Now I began to do beautiful tricks  considering
 the global rotations
\def\a{\alpha}
\def\b{\beta}

                $$
         \cases
          {
          x\rightarrow  x\cos\a+z\sin\a\cr
          z\rightarrow -x\sin\a+z\cos\a\cr
                      }\qquad
         \cases
          {
          x\rightarrow  x\cos\b+y\sin\b\cr
          y\rightarrow -x\sin\b+z\cos\b\cr
                      }\qquad
             $$

Where the angles $\a,\b$ are defined by the values
 of $z_t,y_t$ and I came to the (3).
ITo be fair I have to tell you that from the very beginning
 I noticed that determinant of the matrix (4) is the density
 and invariant density and I couldnot see nothing more--
 and everything have to be combined via this determinant?.
but I did not understand how the invariant combinations of minors arise.

 {\bf Conjecture.} Looking on the sequence (5)  I did the following
  conjecture:
     In the $d$-dimensional space one can consider the sequence
       $\{A_1,A_2,\dots,A_d\}$ where
                  $$
   A_1=\Delta_1,  A_2={\Delta_2\over A_1^2},
          A_3={\Delta_3\over A_2^2\cdot A_1^3},
          A_4={\Delta_4\over A_3^2\cdot A_2^3\cdot A_1^4},
                       \dots
         A_d={\Delta_d\over A_{d-1}^2\cdot A_{d-2}^3\cdot
              A_{d-3}^4\dots A_1^d}
                         $$
   In general the recurrent formula maybe is
                                $$
                        A_{n+1}=
      {\Delta_{n+1}\over\prod_{k=1}^{k=n} A_k^{n+2-k}}
              $$
 I have not any proof yet. Of course
   in the way how I constructed the (3)
  I can consider the sequence of the maps
     from the curve $L$ on the $S^{d-1}$ using the unit vectors
 but I affraid now to consider this after my calculations
for $S^2$...

There are the fo following hints that
    this sequence may be is true:

  1) in the first three terms this is right

  2) the dimension is always the same $[{1\over sec}]$

 Sure this has to come from 19 century. Indeed in classical mathematics
 for the curve in $d$-dimensional space there are the sequence of Frenet
  curvatures $\{k_1,...,k_{d-1}\}$. The curvature $k_1$ is proportional
 to $A_2$:
               $$
    k_1={\Delta_2\over\Delta_1^3}={A_2\over\Delta_1}
                $$
  it meausures in $[{1\over cm}]$.
My sequence have to be strictly connected with the sequence
 of curvatures but my goal was to come to divergence-like
 lagrangians. In some sence this means that the last
curvature is proportional to the divergence-like Lagrangian.

  One can consider the $d$-basis in every point of the curve.
The first vector is tangent vector, the first two-dimensional plane
is the plane spanned by the tangent vector and its acceleration
during the motion, e.t.c.--- Hence the movement on the
  $SO(d)$ of the solid particle can be considered.-- I think
it is just what people call rigid particle motion,
 (Armen considered Lagrangians depended on higher curvatures.)
  I think that all these formulaes may be in the other
 interpretation have to be revealed in rigid particle physics.
   In fact I sent the beginning of this text to Armen,
  and still do not know his reaction.
  Do you want to know why and how I spent yesterday
 morning and day on this calculations?
          S Vas butylka!---Togda skazhu!

            Do svidanija O.M.
\bye
