\magnification=1200

\def\R {{\bf R}}
\def\a {\alpha}




   \centerline {\bf Zorn lemma application}

{\it
  We apply Zorn lemma to show that every ideal belongs to maximal ideal.
Can we apply Zorn lemma to show that every family of ideals
has maximal ideal?  Answer is: "No" since positive answer means that ring is 
noetherian. Let us see how conditions of applying Zorn lemma are failed.}


  Let $I$ be an arbitrary ideal in an arbitrary ring $\R$.

   Consider two sets of ideals:

      $$
   M_1=\{\hbox{
set of all finitely generated ideals of ring $\R$ which 
belong to $I$}\}\,
      $$
and
      $$
   M_2=\{\hbox{
set of all ideals of ring $\R$ which possesses the ideal $I$}\}\,,\quad
      $$
 The first set for not-Noetherian rings is not inductive: it may happen
that $M_1$ possesses
well-ordered subset which has not upper bound in $M_1$.
E.g.  consider ring $\R=\R[x_i]$, 
ring of polynomials with real coefficients
which depend on variables 
$\{x_i\}, i=1,2,3,\dots$ 
(every polynomial depend on finite number of variablles). Consider ideals
   $$
  \a_k=\hbox {polynomials which vanish if $x_1=x_2=\dots=x_k=0$}
   $$
and
      $$
   I=\hbox{polynomials which vanish if at least on of variables is 
equal to zero}
      $$

 The ideal $I$ is not finitely generated. Well-ordered 
set of finitely generated
ideals $\a_k$ which belong to $I$ has not upper bound
in the set $M_1$ of finitely generated ideals. We cannot use Zorn lemma
to prove that the family $M_1$ possesses maximal ideal:
the fact that $M_1$ possesses maximal ideal means that $I$ is finitely 
generated. Indeed let $\beta$ be maximal element in $M_1$, then $\beta=I$,
since if $\beta\not=I$, we can expand $\beta$. 
{Recall that the fact that every family of ideals possesses maximal ideal
is nothing but Noether condition.}




  The second set is inductive: every well-ordered subset $N\subseteq M_2$
has upper bound in $M_2$. Indeed consider subset $N$
 of ideals $N=\{\a_\iota\}, N\subseteq M$
such that for arbitrary two ideals $\a_{\iota_1},\a_{\iota_2}\in N$
              $$
        \a_{\iota_1}\subseteq \a_{\iota_2}\,,{\rm or}\,\,
        \a_{\iota_2}\subseteq \a_{\iota_2}\,.
              $$
(This is well-ordered subset in poset\footnote{$^{1}$}
  {`poset' means {\it partially ordered set}} $M$ of ideals.)

  One can see that $N$ has upper bound in $M$
(not obligatory in $N$ itself!). Indeed consider
union of all ideals in $N$:
       $$
      T_N=\sum_\iota \a
       $$
All ideals $\a$ do not possess $1$,hence
$T\not=\R$ since it does not possess $1$ and $T\R\in T$. Hence
$T$ is ideal. We see that every well-ordered subset in $M_2$
has upper bond. Hence by Zorn lemma we come to the existence
of maximal element in $M_2$. In particular we proved very imprortant 

{\it Proposition} Every ideal is contained in maximal ideal.

\bigskip


 {\bf Resum\'e}.  
  Statement about existence of maximal ideals (in a given family) 
sometimes
follows just from  Zorn lemma, sometimes follows from noetherian
property.

Consider another good example.


{\it Proposition } If $x$ is not nilpotent element of ring $\R$,
then there exist prime ideal which does not possess this element.


Now present the proof based on Zorn lemma.

{\sl Proof}. Consider the set
       $$
M=\hbox {set of all ideals which do not contain
any  power of $x$}
       $$
$M$ is not empty, since $\{0\}\in M$.

  For noetherian ring everything is done:
every set of ideals in noetherian ring possesses  maximal ideal
 \footnote{$^{3)}$}{Indeed suppose this is wrong. Then we come
to infintie ascending series of ideals} and maximal ideal is automatically
simple.

 Now forget about the condition of being noetherian ring.

The set $M$ due to Zorn lemma possess maximal element
(this maximal element is not necessary maximal ideal!)
since every well-ordered subset in $M$ has upper bond

  Hence we come to the ideal $\beta$ such that
            $$
\forall N\,,\quad x^N\not\in \beta_x
             $$
Prove that $\beta$ is a prime ideal.

The proof follows from the lemma

{\bf Lemma} For an arbitrary $a\not \beta$ there exist $N$
such that
        $$
      x^N\in (a)+\beta
        $$

Indeed let $ab\beta$. Suppose that $a,b\not\in\beta$.
The lemma implies that there exist $M.N$ such that 
$x^M-pa\in\beta, x^N-qb\in \beta$ for some elements $p,q\in \R$.
Hence $x^{N+M}\in\beta$ since $ab\in\beta$,  but $x^{N+M}\not\in \beta$.
   Contradiction.

It remains to prove the lemma. 
Suppose $a\not\in \beta$. Consider subring generated by $\beta$ and $a$.
It is all the ring or the ideal. The condition that $\beta$ is maximal 
ideal which does not possess all the powers of $x$ implies the 
statement of lemma.

\bigskip


{\bf Example} Consider algebra of functions on $\R$, 
how looks ideal $\beta_x$. $\beta_x=(x-a)$, where $a\not=0$
 maixmal ideals are points of $\R$









\bye








\bye

