%From ch.duval@gmail.com Wed Dec 15 19:59:09 2010
%Date: Wed, 15 Dec 2010 18:05:57 +0000
%From: Christian Duval <ch.duval@gmail.com>
%To: H Khudaverdian <khudian@manchester.ac.uk>
%Cc: "Christian.Duval@cpt.univ-mrs.fr" <Christian.Duval@cpt.univ-mrs.fr>
%Subject: Re: Letter from Hovik Khudaverdian

Dear Hovhannes,

Sorry for this belated answer to your last email which I read with
great pleasure. The point is that I have, as you can imagine, a tight
teaching timetable before the en of this term.

I am glad that Mrs Bugeau has contacted you and claimed your train
tickets and the like for your reimbursement. Just keep me informed of
what's going on, especially if there is some unwanted delay in the
procedure (which shouldn't take more than 2 months).

As to the topic we discussed in Marseille, I  would of course love to
learn more about your approach regarding this duality
metric/symplectic structure on $\Pi T^*M$ induced by an affine
connection, $\nabla$, on M. Sounds exciting. Affaire \`a suivre \dots

Let me finish by telling you that I have been delighted to make
acquaintance with you, and that the mathematical results you lectured
about in Marseille impressed us quite a lot. Also was the musical
exchange we had extremely exciting.

Till the next!

Yours, friendly,

Christian

%---

On Wed, Dec 15, 2010 at 5:34 PM, H Khudaverdian
<khudian@manchester.ac.uk> wrote:

                      Dear Christian
      I sent facture d'hotel and tickets to
      Julie BUGEAU (Université du Sud Toulon  Var UFR Sciences
      et Techniques)
as she demanded.  (see her letter below).

 I am sorry for may be annuying you: did you receive my
previous letter?

            Thank you very much

                                    Hovik




                                   Dr Hovhannes Khudaverdian
                                  The University of Manchester
                                     School of Mathematics
                                 e-mail:
khudian@manchester.ac.uk
                                      tel. 0161-2008975



On Fri, 10 Dec 2010, H Khudaverdian wrote:


                      Dear Christian,
             Sorry for so long silence.
       My last visit to France, to Lyon and to Marseille
      was extremelly nice: four days which I spent with
      Valya, you and Jean-Phillipe were emotionally very
      nice and mathematically very fruitfull. Honestly it
      happens very rarely when I have had so high level
      of concentration of mathematics and enjoyment
      simulaneously. Thank you very much.

        After my talk in Marseille I was thinking about
      formula which which we discuss after my seminar
      (see below). Unfortunately when I returned to
      Manchester I was dived in every day routine. I
      tried to make something beautiful starting with
      constructions which I thought about in Marseille
      due to the conversations with you. In fact I wanted
      very much to send you the letter not only with "la
      facture d'hôtel et du billet de train Lyon
      Marseille Lyon" but with some beuatiful formulae.
       Helas this did not work.... and i still did not
      send no receipts to France.
       Yesterday I received the letter from Julie BUGEAU
      (Université du Sud Toulon Var UFR Sciences et
      Techniques) (see her letter below).
       Please, tell me do I understand right that I have
      now to send all this stuff (facture de hotel and
      tickets Lyon---Martseille---Lyon) to her?
       In the end of the letter I woud like to tell you
      briefly about the considerations which were
      inspired by our conversation after my talk.
       You remember when after my talk you wrote on the
      blackbopard very beautiful formula for Laplacian on
      $T^*M$ provided a symmetric conenction on $M$?
       What I am thinking about is the following: We can
      consider simultaneously $T^*M$ and $\Pi T^*M$ over
      manifold $M$ (cotangent bundle and contangent
      bundle with reversed parity of fibres, $\Pi$ is
      parity reversing functor). If $\Gamma$ is an
      arbitrary connection on $M$ then symmetric part of
      this connection gives contribution to the
      Riemannian metric on $T^*M$, i.e. to Laplacian on
      $T^*M$ and antisymmetric part of the
      connection---torsion gives contribution to the "odd
      symplectic structure" on $\Pi T^*M$ and to the odd
      Laplacian on the $\Pi T^*M$. (In the case if
      torsion equals to zero we come to standard
      Batalin-Vilkovsiky operator.) There is something
      like "duality" between these both structures. E.g.
      symplectic structures on $T^*M$ and on $\Pi T^*M$
      know only about antisymmetric part of the
      connection $\Gamma$, whilst Riemannian metrics know
      only about symmetric part of connection. The
      principal symbol of Laplacian on $T^*M$ is governed
      by Riemannian metric, i.e. by symmetric part of
      connection and the principal symbol of Laplacian on
      $\Pi T^*M$ is governed by sympelctic structure,
      i.e. by torsion...
       If you want I could send you more detailed file of
      these formulae. This is not big deal, and
      unfortunately I have still no progress but I will
      be  happy to listen any of yours comments and to
      discuss it with you.


        With best regards

                                    Hovik

      PS  By the way  I am reading attentively your
      article
      with Valya and try to make bridges between our
      approach
       (See H.Khudaverdian, T.Voronov "On odd Laplace
      operators. II", Ser. 2, Vol. 212, Amer. Math. Soc.,
      Providence, RI, 2004, 179-205,
      arXiv:math.DG/0212311.) and yours.



\bye



                                        Dr Hovhannes
      Khudaverdian
                                       The University of
      Manchester
                                          School of
      Mathematics
                                      e-mail:
      khudian@manchester.ac.uk
                                           tel.
      0161-2008975



      On Thu, 9 Dec 2010, Julie BUGEAU wrote:

            Bonjour,
            Pouvez-vous signer l'ordre de mission
            en pièce jointe
            s'il vous plait et me le retourné
            accompagné de la facture d'hôtel et du
            billet de train lyon marseille lyon?
            Je vous remercie;
            Cordialement,
            Julie BUGEAU
            Université du Sud Toulon Var
            UFR Sciences et Techniques
            Bât U - BP 20132
            83957 LA GARDE Cédex

            Tel : 0494142075
            Fax : 0494142633


            -------- Message original --------
            Sujet:  Message from Scanner CPT 3eme
            Bat U
            Date :  Thu, 09 Dec 2010 18:17:22 +0100
            De :
            Scan-3eme-BatU-CPT@univ-tln.fr<mailto:Scan-3eme-BatU-CPT@univ-tln.fr>
            Répondre à :
            Scan-3eme-BatU-CPT@univ-tln.fr<mailto:Scan-3eme-BatU-CPT@univ-tln.fr>
            Pour :
             bugeau@univ-tln.fr<mailto:bugeau@univ-tln.fr>





--
Christian Duval
CPT, Case 907
13288 Marseille Cedex 9
France
Tel:  +33491269797
Fax: +33491269553
mailto: duval@cpt.univ-mrs.fr
http://www.cpt.univ-mrs.fr/~duval

