\magnification=1200 %\baselineskip=14pt
\def\vare {\varepsilon}
\def\A {{\bf A}}
\def\t {\tilde}
\def\a {\alpha}
\def\K {{\bf K}}
\def\N {{\bf N}}
\def\V {{\cal V}}
\def\s {{\sigma}}
\def\S {{\Sigma}}
\def\s {{\sigma}}
\def\p{\partial}
\def\vare{{\varepsilon}}
\def\Q {{\bf Q}}
\def\D {{\cal D}}
\def\G {{\Gamma}}
\def\C {{\bf C}}
\def\M {{\cal M}}
\def\Z {{\bf Z}}
\def\U  {{\cal U}}
\def\H {{\cal H}}
\def\R  {{\bf R}}
\def\S  {{\bf S}}
\def\E  {{\bf E}}
\def\l {\lambda}
\def\degree {{\bf {\rm degree}\,\,}}
\def \finish {${\,\,\vrule height1mm depth2mm width 8pt}$}
\def \m {\medskip}
\def\p {\partial}
\def\r {{\bf r}}
\def\v {{\bf v}}
\def\n {{\bf n}}
\def\t {{\bf t}}
\def\b {{\bf b}}
\def\c {{\bf c }}
\def\e{{\bf e}}
\def\ac {{\bf a}}
\def \X   {{\bf X}}
\def \Y   {{\bf Y}}
\def \x   {{\bf x}}
\def \y   {{\bf y}}
\def \G{{\cal G}}

\centerline {Equipartitility  of polygons and poliedra.}


   {\bf Theorem 1} Two polygons of equal area
   are equipartial.

    This means that if polygons $\Pi_1$ and $\Pi_2$ have the same area then one can cut
     the polygon $\Pi_1$ on polygons $\pi_1,dots,\pi_k$ and
     polygon  $\Pi_2$ on polygons $\pi'_1,dots,\pi'_k$ such that polygons $\pi_k$
     are equal to polygons $\pi'_k$: $\pi_1=\pi'_1$,  $\pi_1=\pi'_1$, $\pi_2=\pi'_2,\dots, \pi_k=\pi'_k$.





   I give two hints to prove it.

\m

   {\it Hint 1. This was proved by amateur mathematicien.  This means
    that you can prove it!}

    Hint 2. The proof immediately follows from the lemma.

\m

   {\bf Lemma}  Let $S_1, S_2$ be the triangle and the rectangle respectively such that they have the same area
      and one of the side of triangle $S_1$ coincides with one of the sides of the rectanlge  $S_2$.
     Then the trianlge $S_1$ is equipartial with the rectangle $S_2$.

     {\it Proof}: Evident
  $$ $$

   {\bf Theorem 2} The cub and thetraedre of the same volume are not equipartial.

  It is the one of the Hilbert's problem.

  The meaning of this theorems is following:  We know that area of the triangle is equal to
              $$
              S={ah\over 2}
              \eqno (1)
                $$
  where $h$--is the length of the altitude on the side $a$.

   The volume  of thetraedre is equal to
           $${SH\over 3}\,,
           \eqno (2)
           $$
   where $S$ is the area of the base and
    $h$ is the length of the altitude on the base.

 The Theorem2  means that one cannot escape the Analysis to define the volume of thetraedre.
 (The Theorem 1 claims that one comes to the equation(1) jsut by cutting rectangle)



   Few weeks ago I heard about wonderful proof of the second Theorem. Here it is:


  Consider cube with edge equal to $1$ and regular thetraedre of the same volume.
  Let $\theta$  be an angle between sides of the tetraedre.
  One can see that $\theta\over \pi$ is irrational number.
(I think this follows form the fact that $\cos \theta=1/3$).

  For every polyedron $C$ consider the function
              $$
           P_C=\sum_i l_i F(\varphi_i),\eqno (3)
              $$
where $\{l_i\}$ are edges of the polyedron, $\varphi_i$ is the angle between sides adjusted to the edge $l_i$ugol mezhdu
and $F(\varphi)$--a real valued function (v etoj funktsiji i vsja solj!).
The summation goes over all edges $l_i$.

Now the most interesting part:
 Consider an additive function $F$ on $\R$:
$F(a+b)=F(a)+F(b)$, i.e. the linear function on the real numbers,
considered as a vector space over rational numbers, such that
               $$
             F(\pi/2)=1, F(\theta)=0.
             \eqno (4)
               $$
 To construct this function we need Hamel basis.
  This function is not linear in common sence, i.e. it is not {\it continuous} function.

 Now the proof is an one line:

The function $P_C=\sum_i l_i F(\varphi_i)$ defined by (1,2)
is equal to $8$ if $C$ is the cube of volume $1$
and it is equal to $0$ if $C$ is regular thetraedre.

On the other hand the function  $P_C$ does not change under cutting of polyedron because
the function $F$ is additive function of the angles. Contradiction.\finish



  I enjoyed so much this proof, but something is woryying:
  we use Choice Axiom for constructing additve not continuous function $F$ on all real numbers.
    Do we really need it?

  $$ $$

  I try to give the modify this proof escaping the using of choice axiom.

\m


  Suppose one can cut cube on polyedra $\gamma_1,\dots,\gamma_k$
  such that after putting with each other we come to thetraedre.
   Consider the finite set of angles $\{\varphi_1,\dots,\varphi_N\}$
   which arise during cuttings.

   Let $V$ be the linear space spanned by the numbers $\{\varphi_1,\dots,\varphi_N\}$ with rational coefficients:
              $$
      V=\{a_1\varphi_1+\dots+a_N\varphi_N,\,\,\,{\rm where}\,\,\, a_1,\dots,a_n\in \Q\}
              $$
 $V$ is finite-dimensional vector space (over $\Q$) with dimension $\leq N$.

 Let $F$ be a linear function on $V$ which obeys the condition (4). One does not need choice axiom to
 construct this function (in spite of the fact that a function $F$ is not defined uniquely),
 since


 Nepravda li shto ot ispoljzovanija aksiomy vybora mozhno izbavitjsja
sledujushim rassuzdenijem:
   Pri kazhdoj  konkretnoj protsedure  razrezanij i skleivanij,
privodiashej kub v tetraedr, voznikajet lishj konechnaja posledovateljnosj
uglov. Poetomu shtob pridti k protivorechiju (dlia kazhdogo konkretnoj
protsedury) dostatochno postrojitj linejnuju funktsiju $F$, kotoraja
udovletvoriajet usloviju (2) i opredelena na ratsionlajnoj obolochke
konechnoj poseldovateljnosti uglov,
    voznikajushikh pri dannoj protsedure.

 Praviljno li ja ponimju shto kakby vsjo v poriadke i bez aksiomy vybora?
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

   Krassivo nepravda li?







    \bye


\begin{document}

V$ is finite-dimensional vector space (over $\Q$) with dimension $\leq N
\end{document}
