\magnification=1200


  \centerline {\bf Symmetries}


\bigskip

{\it Symmetries of polygons}
\medskip


{\bf Exercise 1} Calculate multiplication table for $D_3$.



\medskip

 {\bf Exercise 2}  Consider subsets $\{I, S,S^2\}$, $\{I,R_{AB}\}$, $\{S, R_{AB}\}$ in the group $D_3$.
 Are these subsets closed under multiplication?



\medskip

 {\bf Exercise 3$^*$} Describe all subgroups in $D_3$.




\medskip

{\bf Exercise 4}  Consider square  $ABCD$. Find all
symmetry transformations.

 \medskip

   {\bf Exercise 5} How many permutations one can do if we have four elements $\{A,B,C,D\}$ or say
     $\{pen, pencil, chair, table\}$


\medskip

  Is it TRUE or FALSE:  Any permutation of letters $\{A,B,C,D\}$ is a symmetry transformation?

\medskip

 {\bf Exercise 6} Describe symmetries of regular tetrahedra


 \bigskip
{\it Symmetries in algebra}

\medskip

  {\bf Exercise 1}
    Find roots $x_1=A,x_2=B$ of polynomial $x^2-2x-1$.  Are they rational?

    Show that $A^2+B^2$ is rational.

    Is it true that $A^{2007}+B^{2007}$ is rational?

\medskip


    {\bf Exercise 2} Let $A,B$ be roots of polynomial $x^2+x-1$.
     Using Vieta Theorem calculate $a_n=A^n+B^n$ for $n=1,2,3,\dots$ without claculating explicitly roots $A$ and $B$ of the polynomial
    $x^2-x+1$.

\medskip



    {\bf Exercise 3} (Roots of quadratic equation via Vieta theorem)

     Let $A, B$ be roots of quadratic polynomial $x^2+px+q$.
  Using Vieta Theorem calculate $A+B$ and $A-B=\pm \sqrt {(A-B)^2}=....$

  Using these results calulate roots $A$ and $B$.



\medskip

{\bf Exercise 4}. Let $x_1=A,x_2=B$ be roots of  equation $x^2+x-10$.
Then according to Vieta Theorem $A^2+B^2=(A+B)^2-2AB=1^2-2\cdot 10=-19$ is negative number.
 What is wrong here?

\medskip

{\bf Exercise 5}  Let $A,B,C$ be roots of cubic equation $x^3-x-0.1=0$.
Using Vieta Theorem calculate $A+B+C$, $A^2+B^2+C^2$ and $A^3+B^3+C^3$
\medskip

{\bf Exercise 6}

a) Show that numbers $\{2\cos 20^\circ, -2\cos 40^\circ, -2\cos 80^\circ\}$ are roots of the cubic equation
  $x^3-3x-1=0$.
{\it Hint: You may use the formula that $\cos 3\theta=4\cos^3\theta-3cos\theta$}

\smallskip

b) Using Vieta Theorem show that $\cos 20^\circ \cos 40^\circ \cos 80^\circ={1\over 8}$.

  c) Using Vieta Theorem calculate  $\cos^3 20^\circ -\cos^3 40^\circ- \cos^3 80^\circ$




\bye
