\magnification=1200

\def\e{{\bf e}}
\def\f{{\bf f}}
\def\y{{\bf y}}


       \centerline {\bf Lie Theorem of existence and uniqueness 
                  of highest vector}


     \centerline {\bf Important lemma}


   Let $L=H+X_+$ be Lie algebra of upper triangular $n\times n$  matrices.
   ($H$ is Lie algebra of diagonal matrices, $X_+$ is sublagebra if 
strictly upper trinagular matirces.)

   {\bf Lemma} Irreducible representation of $L$ in complex space $V$
is one-dimensional.



    Proof.

Let $V$ is irreducible representation of Lie algebra $L=H+X_+$.

  Let $V_0$ be a subspace of $V$ such that it is irreducible with respect
to subalgebra $X_+$. Then it is easy to see that $V_0$ is one-dimensional.

The proof easy follows from Shur Lemma and commutation relations.
Indeed first note that operators $\{E_{1i}\}$ belong to the centre of the 
algebra $X_+$. Hence all these operators are proprotional to identity
operator on $V_0$. On the other hand all the operators $\{E_{1i}\}$ 
except the operator $E_{12}$ are traceless on $V_0$ since they belong 
to subalgera $[X_+,X_+]$ (they are expressed via commutators).
Hence we proved that for $i=3,\dots,n$ all the
operators $E_{1i}=0$ on $V_0$ and $E_{12}$ is a scalar operator.
 using this fact and Shur lemma we immediately come to the result
that   


  Now take an arbitrary such a subspace $V_0$ in $V$
which  is irreducible one-dimensional representation of $X_+$.
  For every $x\in X_+$ denote by $\lambda(x)$ the eigenvalue
of $x$ on $V_0$:
            $$
       x(\e_0)=\lambda (x)\e_0
            $$
  where $\e_0\in V_0$.


  Consider the flag of subspaces
              $$
         V_0\subset V_1\subset V_2\subset\dots \subset V_k 
           $$
such that all the factor spaces $V_{r+1}\backslash V_r$
 have weightfunction  $\lambda(x)$ for elements of the algebra  $X_+$: 
if $\{\e_0,e_1,\dots,\e_k\}$ s a basis in $V_k$ adjusted to the flag
  ($\e_s\in V_s$) then     
        $$
 \hbox {for every $x\in X_+$}\qquad  
    x(\e_s)=\lambda (x) \e_s+\dots
              $$      
where dots are vectors belonging to $V_{s-1}$.

Suppose now that $V_k$ is maximal flag with weight function $\lambda(x)$
in the vector space $V$,
i.e. for every non-zero vector $\f\in V$ such that $\f\not\in V_k$
there exists an element  $x\in X_+$ such that $x(\f)-\lambda(x)x$
does not belong to $V_k$ also.

By definition $V_k$ is invariant with respect to subalgebra $X_+$.
Show that this is invariant with respect to the whole algebra $L$ also.

Indeed consider a vector $\y=E_{pp}\e$, where $\e\in V_k$ ($E_{pp}\in H$)
Suppose that $\y\not\in V_k$. From commutation relations
            $$
[E_{pp}, E_{ij}]=\delta_{pi}E_{pj}-\delta_{jp}E_{ip}
         \eqno(1)
            $$
it follows that
            $$
       x(\y)=\lambda(x)\y+\dots
          $$
where we denote by dots vector in $V_k$. Hence the flag $V_k$ is not maximal.
Contradiction. So we proved that the flag space $V_k$ is invariant with respect to the whole alge
bra $L$. Hence the flag space $V_k$ coincides with 
the space $V$, $V=V_k$.  Consider again an arbitrary element $x\in X_+$.
 The fact that  $\lambda(x)$ is the weight function of the flag, i.e.
       $$
  x(\e_k)=\lambda(x)+\hbox {vector in $V_{k-1}$}
      $$
means that trace of an arbitrary operator $x$ is equal to $N\lambda(x)$,
where $N$ is dimension of the space $V=V_k$. On the other hand due to 
 commutation relations (*) trace of an arbitrary operator $x\in X_+$
is equal to $0$. Hence we have proved that $\lambda(x)=0$ (for $x\in X_+)$.
Now we see that the action of Lie algebra $L=H+X_+$ on the space 
$V$ is reduced to the action of abelian Lie algebra $H$. Hence
$V$ is $1$-dimensional. Lemma is proved.


$$ $$
  Now we are ready to prove the Theorem.

  A finite dimensional representation of the algebra $gl(n,C)$
or $sl(n,C)$ has a (up to a coefficient) highest vector.
It is a unique (up to a multiplier) if representation is irreducible.


Indeed consider in $W$ a subspace $V$ which is irreducible representation
of the subalgebra $L=H+X_+$ (for $sl(n,C)$ we consider traceless 
diagoanl matrices). Due to the lemma this is one-dimensional vector space.
It is spanned by a vector $\e$. This vector is highest vector.


Let $W$ be finite-dimensional 




\bye        
