
                  \centerline {\bf 1.5 $\pi(n)$ for prime numbers}

 {\bf Theorem}

           $\pi(n)\approx {n\over \log n}$



Two years ago I obtained  empirical proof of this Theorem
 considering $N!$ for large $N$ and using the fact that
 for every primary $p$ $N!$ is divisible on $p^r$
 where
   $$
  r\approx \left[N\over p\right]+\left[N\over p^2\right]+\dots
  \approx {Np\over p-1}
         $$
Then considering integrals instead series we come to the "proof".
It is very rough but I guess that the classic proof is founded on
this remark.

\smallskip

Few days ago in the article about Erdos (The Mathematical Intelligence,
{\bf 19}, 2, pp.38---48) I found the marvelous proof that blongs to Erdos.

I will try to write here about his proof.

  Denote by $\Pi (m,n)$ the product of primary numbers
  that belong to the segment $[m,n]$ ($m<n$):
                      $$
                      \Pi(m,n)=\prod_{m\leq p\leq n} p
                      \quad \hbox {where $p$ are prime numbers}
                      $$

  {\bf Lemma 0}.   $C^m_n\leq 2^{n-1}$ (at least for odd n)

\medskip

  {\bf Lemma 1}. If $2m>n-1$ then $\Pi(m,n)\leq C^m_n$.

\medskip

  {\bf Lemma 2}. $\Pi (1,n)<4^n$!!!.

 From this Lemma it obviously follows that
               $$
          \pi(n)<{n\over \log_4 n}
          $$

\medskip

  Lemma 2 belongs to Erdos and
  {\it a brilliant idea, a suprising step
     to prove it is} (from Math. Intell.) is the  content of Lemma 1.

 the statement of Lemma 0 is trivial. If $n$ is odd then
                  $$
   C^m_n={1\over 2}(C^m_n+C^{n-m}_n)<{1\over 2}\sum C^k_n=
         {1\over 2}2^n=2^{n-1}
                   $$
 The statement of lemma 2 is obvious as soon as it is formulated!

 Now prove Lemma 2 by induction. For n=2 it is valid.

 Let $\Pi (1,k)<4^k$ for all $k<n$. Prove for $k=n$.
 If $n$ is even then $\Pi(1,n)=\Pi(1,n-1)$.
 If $n=2q+1$ is odd then using inductive hypothesis and Lemma 1
  we come to
 $$
 \Pi(1,n)=\Pi(1,q+1)\cdot\Pi(q+2,2q-1)<4^{q+1}C^{q+2}_{2q-1}<
     4^{q+1}2^{2q-2}=4^{2q}\,.$$

 \bye
